The Design of an Aliasing-Tolerant Parallel Programming Model

### 2012-11-09 doc/types-pldi2013/langdef.tex
langdef

Figure~\ref{fig:langdef} defines Core Legion, a subset of the full Legion language
that still illustrates all of the important issues.  
%Types include booleans, integers, tuples, and pointers.  
Pointers are annotated with a list of logical regions---a non-null pointer must point to a 
location that is contained in at least one of the regions. There is a special
{\em coloring} type, used to specify partitions of regions into subregions.
%Functions in Legion are named, accept one or more arguments and
%return a value of some type.  
Functions declare the region
access privileges and coherence they require.  Some function instances are {\em tasks}; a function may be executed
within the current task or run in parallel as a separate (sub)task---this decision is made
by the Legion task scheduler.  Function types are
universally quantified over all region names appearing in the type.
The other important type is a {\em region relationship}, an instance of which captures a value and
one or more regions satisfying associated constraints.  A region relation instance
can be written into the heap and later read and
unpacked, giving new local names to the regions contained in
the instance.

To achieve high performance in the presence of aliasing, the Legion type system
leverages both static and dynamic checks, resulting in interesting extensions beyond
a basic expression language to both the type system and the operational semantics.
The following two subsections illustrate these decisions and how they have been
chosen to preserve a straightforward proof of type soundness.

@@

To determine whether an interleaving of two or more memory traces is valid, we consider three sets of addresses:
\begin{itemize}
\item exclusive locations ($l \in L_{excl}$) are those which have at least one access in exclusive mode in the traces
and are not in the clobber set.  For these locations, we require sequential execution semantics - all
reads to these locations see the effect of previous writes and reductions, and the resulting state of the store is
as if all writes and reductions were applied from each trace in order.
\item atomic locations ($l \in L_{atomic}$) are those which have at least one access in atomic mode in the traces and
are in neither $L_{excl}$ nor the clobber set.  For these locations, we allow permutations of the original
subexpression trace order.
\item for locations that have only access in {\em simult} mode or are in the clobber set, no constraints are
enforced.  The validity of the interleaving of these accesses is performed by some enclosing expression.
\end{itemize}

>>>>

Figure~\ref{fig:langdef} defines Core Legion, a subset of the full Legion language
that illustrates the important issues.  
%Types include booleans, integers, tuples, and pointers.  
Pointers are annotated with a list of logical regions---a non-null pointer must point to a 
location that is contained in at least one of the regions. There is a special
{\em coloring} type, used to specify partitions of regions into subregions.
%Functions in Legion are named, accept one or more arguments and
%return a value of some type.  
Functions declare the region
access privileges and coherence they require.  Some function instances are {\em tasks}; a function may be executed
within the current task or run in parallel as a separate (sub)task---this decision is made
by the Legion task scheduler.  Function types are
universally quantified over all region names appearing in the type.
The other important type is a {\em region relationship}, an instance of which captures a value and
one or more regions satisfying associated constraints.  A region relation instance
can be written into the heap and later read and
unpacked, giving new local names to the regions contained in
the instance.

To achieve high performance in the presence of aliasing, the Legion type system
leverages both static and dynamic checks, resulting in interesting extensions beyond
a basic expression language to both the type system and the operational semantics.
The following two subsections illustrate these decisions and how they have been
chosen to allow a straightforward proof of type soundness.

@@

To determine whether an interleaving of two or more memory traces is valid, we consider three sets of addresses:
\begin{itemize}
\item exclusive locations ($l \in L_{excl}$) are those which have at least one access in exclusive mode in the traces
and are not in the clobber set.  For these locations, we require sequential execution semantics - all
reads to these locations see the effect of previous writes and reductions, and the resulting state of the store is
as if all writes and reductions were applied from each trace in order.
\item atomic locations ($l \in L_{atomic}$) are those which have at least one access in atomic mode in the traces and
are in neither $L_{excl}$ nor the clobber set.  For these locations, we allow permutations of the original
subexpression trace order.
\item for locations with only access in {\em simult} mode or in the clobber set, no constraints are
enforced.  The validity of the interleaving of these accesses is performed by some enclosing expression.
\end{itemize}
